---
title: "Activity 6"
output: html_notebook
---

# Web Reference

https://beta.rstudioconnect.com/content/1518/notebook-classification.html

# Import Libraries

```{r}
library(modelr)
library(dplyr)
library(ggplot2)
library(sparklyr)

set.seed(77)
```

# Helper Functions

```{r}
#' ROC curve code
#'
#' Based on algo 1 page 866, Fawcett2005
#'
#' @param L observations 
#' @param f, predicted prob.
#'
#' @return points in ROC space and score
get_roc <- function(L, f) {
  # Calculate P and N
  P <- sum(L==1)
  N <- sum(L==0)
  # Order the observations by prediction
  df  <- tibble(L, f)
  df <- df %>% arrange(desc(f))
  # Set TP and FP to zero
  TP <- 0
  FP <- 0
  # Set up matrix for results
  R <- NULL
  # Set previous f
  f_prev <- -Inf
  # set counter
  i <- 1
  while(i <= length(df$L)){
    if( df$f[i] != f_prev){
      R <- rbind(R, c(FP/N, TP/P, df$f[i]))
      f_prev <- df$f[i]
    }
    if(df$L[i] == 1){
      TP <- TP + 1
    } else {
      FP <- FP + 1
    }
    i <- i + 1
  }
  R <- rbind(R, c(FP/N, TP/P, f_prev))
  R <- data.frame(R)
  colnames(R) <- c("FPR","TPR", "Score")
  return(R)
}
```


# Connect to Spark

```{r}
sc <- spark_connect(master = "local")
```

# Data Preperation

```{r}
height <- modelr::heights
height <- height[-c(721,973,1611,1641,1866,2828,3122,6178,6319,6794),]
height$marital <- as.numeric(recode(height$marital, "divorced"="0", "married"="1", "separated"="2", "single"="3", widowed="4"))
height$sex <- ifelse(height$sex=="male",0,1)
```

## Copy Data to Spark

```{r}
height_sc <- copy_to(sc, height)
```

## Create Training and Testing partitions

```{r}
partitions <- height_sc %>% 
  sdf_random_split(training = 0.8, test = 0.2)

train_tbl <- partitions$train
test_tbl <- partitions$test
```

# Questions

## Question 1

How many observations does the testing set contain?

```{r}
count(test_tbl)
```

## Question 2

What is the value of the coefficient associated with height in the sex ~ income + height model?

### Predictions

```{r}
ml_log <- test_tbl %>% 
  ml_logistic_regression(sex ~ income + height)

summary(ml_log)
```
#### ROC Curve

```{r}
pred_lr <- ml_predict(ml_log, test_tbl) %>% collect
pred_lr$p1 <- unlist(pred_lr$probability)[ c(FALSE,TRUE) ]

ROC_lr <- get_roc(L = pred_lr$sex, f = pred_lr$p1)
ggplot(ROC_lr, aes(x = FPR, y = TPR)) + geom_line(aes(col = "LR prediction"))
```

## Question 3

### Predictions

```{r}
ml_log2 <- test_tbl %>% 
  ml_logistic_regression(sex ~ income + height + education)

summary(ml_log2)
```
#### ROC Curve

```{r}
pred_lr2 <- ml_predict(ml_log2, test_tbl) %>% collect
pred_lr2$p1 <- unlist(pred_lr2$probability)[ c(FALSE,TRUE) ]

ROC_lr2 <- get_roc(L = pred_lr2$sex, f = pred_lr2$p1)
ggplot(ROC_lr2, aes(x = FPR, y = TPR)) + geom_line(aes(col = "LR prediction"))
```

```{r}
ggplot(ROC_lr, aes(x = FPR, y = TPR)) + 
  geom_line(aes(col = "LR1 prediction")) + 
  geom_line(data = ROC_lr2, aes(col = "LR2 prediction"))
```

