---
title: "Activity 5"
output: html_notebook
---

```{r}
install.packages("rattle.data")
```

```{r}
library(dplyr)
library(sparklyr)
sc <- spark_connect(master = "local")
```
```{r}
wine <- rattle.data::wine
wine[-1] <- scale(wine[-1])
```

Copy the wine data into Spark.

```{r}
wine_tbl <- copy_to(sc, wine, "wine", overwrite=TRUE)
```

Next, fit a PCA model to the wine data, making sure to remove the Type variable before you do so. 

```{r}
pca_model <- tbl(sc, "wine") %>%
  select(-Type) %>% na.omit() %>%
  ml_pca()
```

```{r}
print(pca_model)
```
You’re going to try clustering your data using only two variables - Nonflavanoids and Hue. Because there are three types of wine in our dataset, it makes sense to specify three clusters.

```{r}
kmeans_model <- wine_tbl %>%
  ml_kmeans(formula= ~ Nonflavanoids + Hue, k=3)
print(kmeans_model)
```
What is the within set sum of squared errors for your k-means clustering?
>No idea how the fuck to do this, it is not the below.

```{r}
pred_kmeans <- ml_predict(kmeans_model, wine_tbl)
ml_clustering_evaluator(pred_kmeans)
```



