---
title: "Major Assignment 01 : Part 5"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(nycflights13)

set.seed(19)

flights <- nycflights13::flights
flights$early <- as.numeric(flights$arr_delay < 0)

early_flights <- flights %>% 
  filter(early == 1) %>%
  summarise(total=sum(early))

early_total <- early_flights[[1]]
early_total
```
2.	Build a model to predict arrival delay using origin, departure delay, flight time, carrier, distance travelled, year, month, day and hour. This is a "full" model compared to those in Parts 1-4 of this assignment. 


```{r}
flights.model <- lm(arr_delay ~ origin + dep_delay + air_time + carrier + distance + year + month + day + hour, flights)
summary(flights.model)
```

```{r}
broom::tidy(flights.model)
```

3.	Use backward elimination starting from this model to create a better linear regression model to predict arrival delay. 

```{r}
lm.full <- flights.model
lm.null <- lm(arr_delay ~ 1, data=flights)

step(lm.full, scope=list(lower=lm.null), direction="backward")
```

4.	Apply 10-fold cross-validation, to the model. 

```{r}
cv <- crossv_kfold(flights, k=10)

models <- 
  map(cv$train, ~lm(arr_delay ~ origin + dep_delay + air_time + carrier + distance + month + hour, data=.))

models[[6]]
```

5.	Compute the average mean-square prediction error for arrival delay using this model. 

```{r}
get_pred <- function(model, test_data) {
  data <- as.data.frame(test_data)
  pred <- add_predictions(data, model)
  return(pred)
}

predictions <- map2_df(models, cv$test, get_pred, .id = "Run")
#predictions

MSE <- predictions %>%
  drop_na() %>%
  group_by(Run) %>% 
  summarise(MSE = mean( (arr_delay - pred)^2))

mean(MSE$MSE)
```




