---
title: "Major Assignment 2 - Part 5"
output: html_notebook
---

https://www.r-bloggers.com/2018/12/an-introduction-to-h2o-using-r/

```{r}
# this had to be upgraded
#install.packages("data.table")
#install.packages("bit64")
#install.packages("arulesViz")
```


# Import Libraries

```{r}
library(nycflights13)
library(h2o)
```

# Initiate H2O

```{r}
h2o.init()
```

2.	Load the nycflights13 data as a data frame.

```{r}
flights <- nycflights13::flights
```

3.	Write the data out to a .csv file and read it back again (this solves a data format issue). 

```{r}
write.csv(flights,"flights.csv", row.names=FALSE)
flights <- read.csv("flights.csv")
```

4.	Create a H2O data frame from the data. 

```{r}
flights.hex <- as.h2o(flights)
h2o.describe(flights.hex)
```
5.	Create a new categorical variable in the frame called late_arrival that is 1 if arr_delay is greater than or equal to 30.

```{r}
flights.hex$late_arrival <- as.factor(flights.hex$arr_delay >= 30)
h2o.describe(flights.hex)
```

6.	Create a data split of the H2O frame with 80% of the data being used for training (flight_train) and 20% used for testing (flight_test), ensuring repeatability by setting the seed for the random sampling of the split to be 42. 

```{r}
splits <- h2o.splitFrame(data = flights.hex,
                         ratios = c(0.8),
                         seed = 42)

flight_train <- splits[[1]]
flight_test <- splits[[2]]

print(paste0("Number of rows in train set: ", h2o.nrow(flight_train)))
print(paste0("Number of rows in test set: ", h2o.nrow(flight_test)))
```

7.	Create a predictors set predictors from sched_dep_time, dep_delay, air_time, and distance. 

```{r}
predictors = c("sched_dep_time", "dep_delay", "air_time", "distance")
```

8.	Create a deep learning H20 model using the defaults except for:

-	the predictors being predictors,
-	the response variable being late_arrival,
-	the training data being flight_train, and
-	the results forced to be reproducible with seed 42.

Run several times to check your results are reproducible.

```{r}
flights.dl <- h2o.deeplearning(x=predictors, y=c("late_arrival"), training_frame=flight_train, seed=42)
summary(flights.dl)
```

9.	Plot an ROC curve for your model using the flight_test data. 

```{r}
flights.perf <- h2o.performance(model=flights.dl, newdata=flight_test)
plot(flights.perf, type="roc")
```

10.	Calculate the prediction accuracy produced by your model using h2o.predict on your model and test data flight_test. 


```{r}
pred <- h2o.predict(flights.dl, newdata=flight_test)
```

```{r}
flights.perf
```

11.	Calculate the prediction accuracy that would be yielded, for the test data flight_test, by simply using dep_delay (being larger than 30) alone as a predictor. 

```{r}
flights.hex$late_departure <- as.factor(flights.hex$dep_delay > 30)

splits <- h2o.splitFrame(data = flights.hex,
                         ratios = c(0.8),
                         seed = 42)

flight_train <- splits[[1]]
flight_test <- splits[[2]]

flights.dl <- h2o.deeplearning(x=c("late_departure"), y=c("late_arrival"), training_frame=flight_train, seed=42)

flights.perf <- h2o.performance(model=flights.dl, newdata=flight_test)
flights.perf
```

```{r}
h2o.accuracy(flights.perf)
```










