---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(nycflights13)
```

```{r}
flights <- nycflights13::flights
flights
```

# Part 1

## Scatterplot 

```{r}
ggplot(flights, aes(x=dep_delay, y=arr_delay)) + 
  labs(title="Flight Delays", x="Departure Delay", y="Arrival Delay") +
  geom_point() +
  geom_smooth()
```

## Linear Regression

```{r}
flights.lm <- lm(arr_delay ~ dep_delay, data=flights)
summary(flights.lm)
```

## Residual versus Fitted plot

```{r}
plot(flights.lm, which = 1)
```

```{r}
ggplot(flights,aes(dep_delay)) + geom_bar()
```

## Normal QQ-plot

```{r}
plot(flights.lm, which=2)
```
## Expected Arrival Delay

```{r}
newdata <- data.frame(dep_delay=c(100))
predict(flights.lm, newdata=newdata)
```

# Part 2

## Scatterplot of arr_delay against dep_delay with colours for the different origins

```{r}
ggplot(flights,mapping = aes(x=dep_delay, y=arr_delay)) + 
  labs(title="Flight Delays", x="Departure Delay", y="Arrival Delay") +
  geom_point(aes(color = origin)) + 
  geom_smooth(method = 'lm', mapping = aes(color=origin))
```

## 3.	Fit a linear model with arr_delay as the response variable; dep_delay as a predictor; origin as another predictor; and an interaction term between dep_delay and origin

```{r}
flights.lm2 <- lm(arr_delay ~ dep_delay + origin + dep_delay:origin, data=flights)
summary(flights.lm2)
```

## Coefficients Table

```{r}
summary(flights.lm2)$coefficients
```

## Predictions

```{r}
newdata <- data.frame(dep_delay=c(0,0), origin=c("EWR", "LGA"))
predict(flights.lm2, newdata=newdata)
```
