---
title: "Major Assignment 01 - Part 3"
output: _html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(nycflights13)
```

```{r}
flights <- nycflights13::flights
flights
```

# Part 3

## Nested dataframe with a row for each carrier

```{r}
flights.nested <- flights %>%
  group_by(carrier) %>% 
  nest()

flights.nested
```

## Fit a linear model, for each carrier

```{r}
flights_model <- function(df) {
  lm(arr_delay ~ dep_delay, data=df)
}

flights.model <- flights.nested %>% 
  mutate(model = map(data, flights_model), 
         coef = map(model, broom::tidy))

flights.model
```

```{r}
yv <- flights.model %>%
  filter(carrier == "YV")

yv$data[[1]]
```

## Table of the coefficients for each carrier model

```{r}
flights.coef  <- flights.model %>% unnest(coef)

flights.coef %>%
  filter(carrier == "YV")
```

```{r}
flights.coef  <- flights.model %>% unnest(coef)

flights.coef %>%
  filter(carrier == "UA")
```

```{r}
flights.coef %>% print(n = Inf)
```

# Part 4

## Count the number of delayed flights

Create a new variable called delayed that has the value 1 if the arrival delay is greater than zero, and value 0 if the arrival delay is less than or equal to zero. Count the number of delayed flights. 

```{r}
flights$delayed <- as.numeric(flights$arr_delay > 0)
flights
```

```{r}
flights %>% 
  group_by(delayed) %>%
  summarise(count=n())
```

```{r}
delayed_flights <- flights %>% 
  filter(delayed == 1) %>%
  summarise(total=sum(delayed))

delayed_total <- delayed_flights[[1]]
delayed_total
```

## Logistic Regression Model

Fit a logistic regression model with delayed as the response variable and origin as the predictor.

```{r}
flights.glm <- glm(delayed ~ origin, family=binomial(), flights)
summary(flights.glm)
```

## Produce the coefficients table for the logistic regression

```{r}
flights.glm_coef <- broom::tidy(flights.glm)
flights.glm_coef
```

## Predictions

```{r}
newdata <- data.frame(origin=c("LGA", "JFK", "EWR"))
predict(flights.glm, newdata=newdata, type="response")
```




