---
title: "Major Assignment 2 - Part 2"
output: html_notebook
---

# Web Reference

https://beta.rstudioconnect.com/content/1518/notebook-classification.html

# Import Libraries

```{r}
library(modelr)
library(dplyr)
library(ggplot2)
library(sparklyr)
library(nycflights13)

set.seed(42)
```

# Helper Functions

```{r}
#' ROC curve code
#'
#' Based on algo 1 page 866, Fawcett2005
#'
#' @param L observations 
#' @param f, predicted prob.
#'
#' @return points in ROC space and score
get_roc <- function(L, f) {
  # Calculate P and N
  P <- sum(L==1)
  N <- sum(L==0)
  # Order the observations by prediction
  df  <- tibble(L, f)
  df <- df %>% arrange(desc(f))
  # Set TP and FP to zero
  TP <- 0
  FP <- 0
  # Set up matrix for results
  R <- NULL
  # Set previous f
  f_prev <- -Inf
  # set counter
  i <- 1
  while(i <= length(df$L)){
    if( df$f[i] != f_prev){
      R <- rbind(R, c(FP/N, TP/P, df$f[i]))
      f_prev <- df$f[i]
    }
    if(df$L[i] == 1){
      TP <- TP + 1
    } else {
      FP <- FP + 1
    }
    i <- i + 1
  }
  R <- rbind(R, c(FP/N, TP/P, f_prev))
  R <- data.frame(R)
  colnames(R) <- c("FPR","TPR", "Score")
  return(R)
}
```


# Connect to Spark

```{r}
sc <- spark_connect(master = "local")
```

# Data Preperation

```{r}
flights_tbl <- copy_to(sc, nycflights13::flights, "flights", overwrite=TRUE)

flights_final_tbl <- flights_tbl %>%
  ft_binarizer("arr_delay", "late_arrival", threshold=30)  
```

# Question 1

How many late arrivals are there in the dataset?

```{r}
flights_final_tbl %>% 
  group_by(late_arrival) %>%
  summarise(count=count())
```

# Question 3

What is the coefficient for the carrier UA in the logistic regression model?

## Create Training and Testing partitions

```{r}
partitions <- flights_final_tbl %>%
  na.omit() %>%
  sdf_random_split(training=0.75, test=0.25, seed=42)

train_tbl <- partitions$train
test_tbl <- partitions$test
```

## Use logistic regression to predict late arrivals based on carrier, departure delay, month, and year 

```{r}
ml_log <- test_tbl %>% 
  ml_logistic_regression(late_arrival ~ carrier + dep_delay + month + year)

summary(ml_log)
```
# Question 5

What is the true positive rate when the false positive rate is 0.5 in the model?

## Perform Predictions

```{r}
pred_lr <- ml_predict(ml_log, test_tbl) %>% collect
pred_lr$p1 <- unlist(pred_lr$probability)[ c(FALSE,TRUE) ]
```


## Calculate the ROC values and plot the curve

```{r}
ROC_lr <- get_roc(L = pred_lr$late_arrival, f = pred_lr$p1)
ggplot(ROC_lr, aes(x = FPR, y = TPR)) + geom_line(aes(col = "LR prediction"))
```
What is the true positive rate when the false positive rate is 0.5 in the model?

```{r}
ROC_lr %>%
  filter(FPR >= 0.5)
```

# Question 6

Remove each predictor variable individually to determine which is the best single predictor of late arrivals.

```{r}
calculate_roc <- function(form, data) {
  # create the model
  ml_log <- test_tbl %>% 
    ml_logistic_regression(form)
  
  # perform the predictions
  pred_lr <- ml_predict(ml_log, test_tbl) %>% collect
  pred_lr$p1 <- unlist(pred_lr$probability)[ c(FALSE,TRUE) ]
  
  # calculate the roc values
  return (get_roc(L = pred_lr$late_arrival, f = pred_lr$p1))
}
```

```{r}
roc1 <- calculate_roc(formula(late_arrival ~ carrier + dep_delay + month + year), test_tbl)
roc2 <- calculate_roc(formula(late_arrival ~ dep_delay + month + year), test_tbl)
roc3 <- calculate_roc(formula(late_arrival ~ month + year), test_tbl)
roc4 <- calculate_roc(formula(late_arrival ~ year), test_tbl)


ggplot(roc1, aes(x=FPR, y=TPR)) + 
  geom_line(aes(col="carrier + dep_delay + month + year")) + 
  geom_line(data=roc2, aes(col="dep_delay + month + year")) +
  geom_line(data=roc3, aes(col="month + year")) +
  geom_line(data=roc4, aes(col="year"))
```

```{r}
ggplot(roc1, aes(x=FPR, y=TPR)) + 
  #geom_line(aes(col="carrier + dep_delay + month + year")) + 
  #geom_line(data=roc2, aes(col="late_arrival ~ carrier + dep_delay + month")) +
  geom_line(data=roc3, aes(col="late_arrival ~ carrier + dep_delay")) +
  geom_line(data=roc4, aes(col="late_arrival ~ carrier"))
```




```{r}
ml_log <- test_tbl %>% 
  ml_logistic_regression(late_arrival ~ dep_delay)

pred <- ml_predict(ml_log, test_tbl)
ml_binary_classification_evaluator(pred, metric_name = "areaUnderROC")
```






