---
title: "Activity 08"
output: html_notebook
---

```{r}
small_data <- data.frame( 
                 X = c(1,1,-1,1),
                 Y = c(1,-1,1,1),
                 Z = c(1,1,1,-1)) 
small_data
```


## Part 1

Use the `small_data` dataframe containing the data matrix D to produce a scatterplot of Y against X, Z against X, and Z against Y. When you have done that, answer Question 1.

## Part 2

Calculate the covariance matrix V of the variables X, Y, and Z. When you have done that, answer Questions 2-4.

```{r}
small_data %>% 
  ggplot(aes(x=X, y=Z)) +
  geom_point()
```
```{r}
cov(small_data$X, small_data$Y)
cov(small_data$Z, small_data$X)
cov(small_data$Z, small_data$Z)
```
```{r}
D = as.matrix(small_data)

E = cbind(
  c(-1, 0, 1),
  c(-1, 1, 0),
  c(1, 1, 1))

print("E:")
E

P=D %*% E

print("P:")
P
```

```{r}
data.frame(P)
```
```{r}
data.frame(P) %>%
  ggplot(aes(x=X2, y=X1)) +
  geom_point()
```


## Part 4

Multiply the original data matrix V by E to obtain the matrix of principal components P. When you have done that, answer Questions 5-7.