---
title: "Major Assignment 1"
output: html_notebook
---

# Import Libraries

```{r}
library(tidyverse)
library(nycflights13)
```

# Part 1

## 1.1 Read in the dataset 

```{r}
flights <- nycflights13::flights
flights
View(flights)
```
## 1.2 Produce a table that summarises the origin variable 

```{r}
origin_summary <- flights %>%
  select(origin) %>%
  group_by(origin) %>%
  summarise(n=n()) %>%
  arrange(desc(n))

origin_summary
```
## 1.3 Produce a bar chart of the origin variable

```{r}
flights %>%
  ggplot(aes(origin)) +
  ggtitle("Flight Origin") +
  geom_bar()
```
## 1.4 Calculate the mean and standard deviation of the distance variable

```{r}
summary(flights$distance)
```
## 1.5 Produce a histogram of the distance variable 

```{r}
flights %>%
  ggplot(aes(distance)) +
  ggtitle("Distance between airports, in miles") +
  geom_histogram(bins=30)
```
# Part 2

## 2.1 Produce a scatterplot

```{r}
airtime_distance <- flights %>%
  select(air_time, distance) %>%
  drop_na()

airtime_distance %>%
  select(air_time, distance) %>%
  drop_na() %>%
  ggplot(mapping = aes(x=distance, y=air_time)) +
  ggtitle("Air Time vs Distance") +
  geom_point() + 
  geom_smooth()
```
```{r}
cor(airtime_distance$air_time, airtime_distance$distance)
```
## 2.2 Produce a contingency table of carrier and origin

```{r}
contingency_flights_origin <- table(flights$carrier, flights$origin)
contingency_flights_origin
View(nycflights13::airlines)
```
## 2.3 Produce a conditional table of carrier and origin to find the percentage (%) of American Airlines departing from JFK

```{r}
contingency_flights_origin_prop <- prop.table(contingency_flights_origin)
contingency_flights_origin_prop * 100
```
# 2.4. Produce side-by-side (one next to the other) boxplots of distance for each origin

```{r}
flights %>%
  select(origin, distance) %>%
  drop_na() %>%
  ggplot(aes(distance)) +
  geom_boxplot() +
  ggtitle("Distance Summary by Origin") + 
  coord_flip() +
  facet_wrap(~origin)
```
# Part 3

## 3.1 Calculate how many flights departed from JFK in May 2013

```{r}
flights %>%
  filter(origin == "JFK", month == 5) %>%
  count()
```
## 3.2 Identify the carrier and the airport that had the first flight of 2013

```{r}
flights %>%
  arrange(year, month, day, dep_time) %>%
  left_join(airlines, by="carrier") %>%
  select(carrier, name, origin) %>%
  head(1)
```
## 3.3 Calculate the total metres travelled by flights departing NYC in 2013 (you may assume 1609.34 metres/mile)

```{r}
miles_meters <- 1609.34

miles_travelled <- flights %>%
  select(distance) %>%
  sum()

miles_travelled * miles_meters
```
## 3.4 Calculate the total distance in miles travelled by flights departing from NYC in 2013 for airlines with Inc. in the name 

```{r}
flights %>%
  left_join(airlines, by="carrier") %>%
  select(name, distance) %>%
  filter(str_detect(name, "Inc.")) %>%
  select(distance) %>%
  sum()
```
# Part 4

## 4.1 Calculate how many standard deviations greater than the mean the largest distance in the flights dataset is.

```{r}
flight_distance <- flights %>%
  select(distance) %>%
  drop_na()

# get the summary statistics
distance_mean <- mean(flight_distance$distance)
distance_sd <- sd(flight_distance$distance)
distance_max <- max(flight_distance$distance)

# calculate the number of sd away
(distance_max - distance_mean) / distance_sd
```
# Part 5

## 5.1 Calculate the 95% confidence interval for the mean distance for flights departing NYC

```{r}
# get the total number of flights
distance_n <- length(flight_distance$distance)

# using student t-distribution 95% confidence interval t value
t <- qt(p=0.025, df=distance_n-1, lower.tail=FALSE)

# calculate the lower and upper values
lwr <- distance_mean - t * distance_sd / sqrt(distance_n)
upr <- distance_mean + t * distance_sd / sqrt(distance_n)

c(lwr = lwr, upr = upr)
```
## 5.2 Calculate 95% confidence interval for the mean distance for each carrier

```{r}
flights %>%
  select(carrier, distance)  %>%
  drop_na() %>%
  group_by(carrier) %>%
  summarise(mean=mean(distance), sd=sd(distance), n=n()) %>%
  mutate(t=qt(p=0.025, df=n-1, lower.tail=FALSE)) %>%
  mutate(
    lower=mean - t * sd / sqrt(n),
    upper=mean + t * sd / sqrt(n)
  ) %>%
  left_join(airlines, by="carrier") %>%
  select(carrier, name, mean, upper, lower)
```

